Measuring Hash Trustworthiness via Collision Utility Metrics: Logical Cryptanalysis of MD4

### 2017-12-05 dsn-2018/bare_conf.tex
DSN changes

We define a relation among families of collisions across rounds as follows.
Let $F_1$ and $F_2$ be two different families of collisions. Then we say that
$F_1$ and $F_2$ are \textit{similar}, and notate it $F_1 \precsim F_2$, if
$R(F_1) \leq R(F_2)$ and $F_1 \subseteq F_2$. Note that, when
$R(F_1) = R(F_2)$, this is equivalent to saying that $F_2$ is in some
expansion neighborhood of $F_1$. Further, when $R(F_1) < R(F_2)$ and
$F_1 = F_2$, then we say that $F_2$ is the \textit{trivial extension} of
$F_1$.

We claim that the following statements are true for MD4:
\begin{enumerate}
    \item For every $F \in \mathcal{F}$, there exists $F'$ such that
        $F' \precsim F$ and $R(F') + 4 = R(F)$.
    \item For every $F \in \mathcal{F}$, there exists $F'$ such that
        $F \precsim F'$.
\end{enumerate}

From an intuitive standpoint, what this says is that collisions have
witnesses for their existence in reduced round versions of the function, and
further, that it is sufficient to search reduced round versions to find
candidates for extension into additional rounds. This likely derives from
the iterative construction of MD4 and the particular choice of block schedule.
We present Figure \ref{Fig:PrecSim} as an example of the lattice structure
induced under $\precsim$ for example collision families.

\begin{figure*}
\begin{center}
\includegraphics[width=\textwidth]{Figs/lattice.pdf}
\caption{The lattice structure of $\precsim$.}
\label{Fig:PrecSim}
\end{center}
\end{figure*}

\subsubsection{Collision Class Similarity}

We claim that the aforementioned statements hold when considering individual
collision classes instead of families of collision classes, but with less
likelihood. That is, we can transplant an $n$ round collision into an $n+4$
round and consider a neighborhood around it to find a new collision.

\subsubsection{Second Preimage Similarity}

If the models are constrained for a second preimage, we note that attack again
works, but requires searching wider and wider differential paths in small round
sizes. In particular, despite strong evidence for this pattern holding in
$n \leq 24$ rounds, outside of the collision family $\{0, 16\}$ in both 28 and
32 round MD4--which does not extend any further--we have been unable to find
any other differential path in those rounds by exhaustive brute force search
up to a differential path of length 8. Further searches have been done of
36, 40, 44, and 48 rounds up to a differential path of length 5 with a similar
lack of results. Thus we believe the general technique to hold, but the size
of the search space for a second preimage becomes prohibitively large, and
further, that differential paths become sufficiently long to prohibit easy
finding of second preimages.

\subsubsection{Chosen Prefix Similarity}

Unlike second preimage attacks, chosen prefix attacks have a higher likelihood
of working, and give better results. However, we have found that most chosen
prefixes work in the neighborhoods of existing collisions, and thus are not
convinced that inductively building a chosen prefix collision is a useful
technique by itself.

>>>>

We define a relation among families of collisions across rounds as follows.
Let $F_1$ and $F_2$ be two different families of collisions. We say that
$F_1$ and $F_2$ are \textit{similar} and notate this similarity as $F_1 \precsim F_2$ if
$R(F_1) \leq R(F_2)$ and $F_1 \subseteq F_2$. Note that, when
$R(F_1) = R(F_2)$, this is equivalent to saying that $F_2$ is in some
expansion neighborhood of $F_1$. Further, when $R(F_1) < R(F_2)$ and
$F_1 = F_2$, then we say that $F_2$ is the \textit{trivial extension} of
$F_1$.

We claim that the following statements are true for MD4:
\begin{enumerate}
    \item For every $F \in \mathcal{F}$, there exists $F'$ such that
        $F' \precsim F$ and $R(F') + 4 = R(F)$ due to the entire
        internal state being replaced every four rounds of MD4.
    \item For every $F \in \mathcal{F}$, there exists $F'$ such that
        $F \precsim F'$.
\end{enumerate}

\begin{figure*}
\begin{center}
\includegraphics[width=\textwidth]{Figs/lattice.pdf}
\caption{The lattice structure of $\precsim$ for the collision
  families under X. Wang and Y. Sasaki.}
\label{Fig:PrecSim}
\end{center}
\end{figure*}

From an intuitive standpoint, what this says is that collisions have
witnesses for their existence in reduced round versions of the function, and
further, that it is sufficient to search reduced round versions to find
candidates for extension into additional rounds. This likely derives from
the iterative construction of MD4 and the particular choice of block schedule.
We present Figure \ref{Fig:PrecSim} as an example of the lattice structure
induced under $\precsim$ for example collision families, and
demonstrates the claims from our statements about MD4 for collision
families under X. Wang and Y. Sasaki.

\paragraph{Collision Class Similarity}

We claim that the aforementioned statements hold when considering individual
collision classes instead of families of collision classes, but with less
likelihood which may lead to the requirement of enlarging the
neighborhood distance. That is, we can transplant an $n$ round collision into an $n+4$
round and consider a neighborhood to find new collisions.  Empirically
we have shown this is not only likely, but extant for published
collision classes.

\paragraph{Second Preimage Similarity}

If the models are constrained for a second preimage, we note that attack again
works, but requires searching wider and wider differential paths in small round
sizes. In particular, despite strong evidence for this pattern holding in
$n \leq 24$ rounds, outside of the collision family $\{0, 16\}$ in both 28 and
32 round MD4 --- which does not extend any further --- we have been unable to find
any other differential path in those rounds by exhaustive brute force search
up to a differential path of length 8. Further searches have been done of
36, 40, 44, and 48 rounds up to a differential path of length 5 with a similar
lack of results. Thus we believe the general technique to hold but it
is heuristic in nature and the size
of the search space for a second preimage may become prohibitively
large so as to be infeasible, and
further, that differential paths become sufficiently long to prohibit easy
finding of second preimages.

\paragraph{Chosen Prefix Similarity}

The attack also has been shown empirically to work on chosen prefix
attacks, but once again these attacks are heuristic and not
guaranteed.  Unlike second preimage attacks, chosen prefix attacks
have empirically shown a higher likelihood
of working, and have given better results in practice. We have found that most chosen
prefixes work in the neighborhoods of existing collisions, and thus are not
convinced that inductively building a chosen prefix collision is a useful
technique by itself.
